\documentclass{article}

% Language setting
\usepackage[english]{babel}

% Set page size and margins
\usepackage[letterpaper, top=2cm, bottom=2cm, left=3cm, right=3cm, marginparwidth=1.75cm]{geometry}

% Mathematical packages
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}

% Color package
\usepackage{xcolor}

\title{Problem Set 1 - Graphical Adventures: Navigating the Nodes!}
\author{Shayan Shahrabi}
\date{October 2024}

\begin{document}
\maketitle

\section*{Warm-up Problems}

\subsection*{Warm-Up 1}
Suppose $G$ is a simple, connected, non-complete graph. Prove that there exists
three nodes $u$, $v$ and $w$ where $uv$, $vw$ $\in$ but $uw$ $\notin$ $E(G)$.


\textbf{Solution:}

Prove by contradiction!

\subsection*{Warm-Up 2}
Suppose that in a graph $G$, there are exactly 2 odd vertices. Prove that there must be a path between these 2 odd nodes.


\textbf{Solution:}
 
Prove by contradiction again :)



\subsection*{Warm-Up 3 (a theorem to be exact)}
Suppose $G$ is a directed graph with a set of nodes $v_i$. Prove that 
\[
\sum_{v \in V} d^+(v) = \sum_{v \in V} d^-(v)
\]
where $d^+(v)$ denotes the out-degree of node $v$ and $d^-(v)$ denotes the in-degree of node $v$.


\textbf{Solution: }

What do these sumations, each count? Remember the very first theorem you learned...

\vspace{20pt}
\hrule
\vspace{20pt}

\section*{Problems}

\subsection*{Problem 1: Node Necessity Conundrum}

In a graph $G$, if we remove the node $v$, then $G$ will not be connected. Prove that there exists 2 nodes in $G$ like $a$ and $b$ where all the paths from $a$ to $b$ pass through $v$.


\subsection*{Problem 2: Minimum Length of a Path}
Prove that in a graph $G$, there exists a path of length at least $\delta$.

\subsection*{Problem 3: Really Simple}
Prove that in every graph $G$, the following equation holds:
\begin{equation*}
\delta \le \frac{2m}{n} \le \Delta
\end{equation*}

\subsection*{Problem 4: Triangle Inequality}
Prove that for any 3 nodes $x$, $y$, and $z$ in a graph $G$, the following inequality holds:
\begin{equation*}
d(x,y) + d(y,z) \ge d(x,z)
\end{equation*}

\subsection*{Problem 5: How to Stay Connected?}
Prove that if $G$ is self-complementary, then it is connected.

\subsection*{Problem 6: Edge-Vertex Relationship in Simple Connected Graphs}
Suppose in a simple and connected graph $G$, we have $diam(G) = 2$ and $\Delta = n - 2$. Prove that
\begin{equation*}
m \ge 2n - 4
\end{equation*}
where $n$ is the number of vertices and $m$ is the number of edges in $G$.

\subsection*{Problem 7: Bonus Question}
\textcolor{red}{\textit{Note: Your solution for the bonus question will be reviewed only after you have solved all the previous questions, and I have approved your solution. Moreover, keep in mind that you must communicate your solution in person; otherwise, no points will be awarded!}}

Suppose $V = \{A_1, A_2, A_3, \ldots, A_n\}$ is a subset of $n$ points in space, such that the distance between any two points is at least 1. Prove that at most $3n$ pairs of these points can have distances exactly equal to 1.

\end{document}